\section{Whitney Topologies}

Let $X,Y$ be smooth manifolds. In this section we discuss the concept of $C^k$-convergence for sequences of functions $(f_n) \subseteq C^k(X,Y)$ with respect to the Whitney topolgies. The notation and terminology used will be the same as in \cite{golubitsky2012stable}. First we remark that $X$ or $Y$ is also allowed to have boundary. The definition of the jet bundles $J^k(X,Y)$ and the Whitney topologies works exactly the same (if both have boundary, $J^k(X,Y)$ will be a manifold with corners, so we might want to avoid that). Let us begin by understanding the $C^0$-topology.

\begin{lemma}
    \label{whitney_top:lem:c0_and_exponential_topology_agree}
    Let $X,Y$ be smooth manifolds with $X$ compact. Then the Whitney $C^0$-topology (as described in \cite{golubitsky2012stable}) and the exponential topology (as described in \cite{steimle_topo}) on $C^0(X,Y)$ agree.
\end{lemma}

\begin{proof}
    We identify $J^0(X,Y) \cong X \times Y$. Then by definition the sets of the form 
    \[
    \mathrm{M}(W) := \{f \in C^0(X,Y) \mid j^0f(X) = \mathrm{Graph}(f) \subseteq W\}
    \]
    for $W\subseteq X \times Y$ open are a basis of the $C^0$-topology on $C^0(X,Y)$. To see that this topology agrees with the exponential topology we show that it satisfies the correct universal property. So let $S$ be any topological space. We have to show that a map 
    \[
    F \colon S \to C^0(X,Y)
    \]
    is continuous if and only if 
    \[
    \widetilde{F} \colon S \times X \to Y, \; (s,x) \mapsto F(s)(x)
    \]
    is continuous. First assume $F$ to be continuous. Let $V\subseteq Y$ be open and $(s_0,x_0) \in \widetilde{F}^{-1}(V)$. Define $f_0 := F(s_0) \in C^0(X,Y)$, then 
    \[
    U := f_0^{-1}(V)
    \]
    is open in $X$. and $x_0 \in U$. Choose some smaller open neighbourhood $U_0 \subseteq X$ of $x_0$ with $\overline{U}_0 \subseteq U$ and set 
    \[
    W := U \times V \cup (X\setminus \overline{U}_0) \times Y \subseteq X \times Y.
    \]
    By assumption $S_0 := F^{-1}(\mathrm{M}(W)) \subseteq S$ is open and we have constructed $W$, such that $s_0 \in S_0$. But then 
    \[
    S_0 \times U_0 \subseteq S \times X
    \]
    is an open neighbourhood of $(s_0,x_0)$ and $S_0 \times U_0 \subseteq \widetilde{F}^{-1}(V)$.

    Now assume that $\widetilde{F}$ is continuous. Let $W\subseteq X \times Y$ be open and $s_0 \in F^{-1}(\mathrm{M}(W))$. Again define $f_0 := F(s_0) \in C^0(X,Y)$. Then $\mathrm{Graph}(f_0) \subseteq W$ and around each $(x,f_0(x)) \in \mathrm{Graph}(f_0)$ we can find a product neighbourhood $U_x \times V_x \subseteq W$, where $U_x \subseteq X$ is an open neighbourhood of $x$ and $V_x \subset Y$ is an open neighbourhood of $f_0(x)$. After possibly shrinking $U_x$ we may further assume that $\overline{U}_x \subseteq f_0^{-1}(V_x)$. Since $X$ is compact we can cover $\mathrm{Graph}(f_0)$ with finitely many such neighbourhoods $U_1 \times V_1,\ldots,U_n\times V_n$. Now for each $V_i$ we have that 
    \[
    \{s_0\} \times \overline{U}_i \subseteq \widetilde{F}^{-1}(V_i).
    \]
    But $\widetilde{F}^{-1}(V_i) \subseteq S \times X$ is open by assumption and $\overline{U}_i$ is compact, so we can find open neighbourhoods $S_i \subseteq S$ of $s_0$ with 
    \[
    S_i \times \overline{U}_i \subseteq \widetilde{F}^{-1}(V_i).
    \]
    Set $S_0 := S_1 \cap \ldots \cap S_n$. Then $S_0 \subseteq S$ is an open neighbourhood of $s_0$ and for any $(s,x) \in S_0 \times X$ there exists $U_i$ with $x \in U_i$, hence 
    \[
    F(s)(x) = \widetilde{F}(s,x) \in V_i.
    \]
    It follows that $\mathrm{Graph}(F(s)) \subseteq \bigcup_{i=1}^n U_i \times V_i \subseteq W$ for all $s \in S_0$ and thus $S_0 \subseteq F^{-1}(\mathrm{M}(W))$.
\end{proof}

\begin{remark}
    If $X$ is not compact the $C^0$-topology will in general be finer than the exponential topology. To see this recall that a subbase for the exponential topology on $C^0(X,Y)$ is given by sets of the form
    \[
    F(K,V) := \{f \in C^0(X,Y) \mid f(K) \subseteq V\}
    \]
    for compact sets $K \subseteq X$ and open sets $V \subseteq Y$. Set 
    \[
    W := X \times V \cup (X\setminus K) \times Y.
    \]
    Then $W \subseteq X \times Y$ is open and 
    \[
    F(K,V) = M(W).
    \]
    Before giving an example where the $C^0$-topology is strictly finer, we fix some terminology to differentiate between the two topologies on $C^0(X,Y)$. We shall use the term $C^0_\mathrm{exp}$-convergence to refer to convergence with respect to the exponential topology and reserve the term $C^0$-convergence for convergence with respect to the Whitney $C^0$-topology. Now consider the following example. Define a sequence 
    \[
    f_n := \mathrm{const}_{\frac{1}{n}}  \colon \R \to \R, \; x \mapsto \frac{1}{n}
    \]
    of constant continuous functions. We claim that 
    \[
    f_n \xrightarrow{C^0_\mathrm{exp}} \mathrm{const}_0.
    \]
    To see this let $F(K,V) \subseteq C^0(\R,\R)$ be any open neighbourhood of $\mathrm{const}_0$ of the form described above. Then $0 \in V$ and so for $n$ sufficiently large $\frac{1}{n} \in V$, which is equivalent to $f_n \in F(K,V)$. Since the sets of the form $F(K,V)$ are a subbase for the exponential topology, this shows the claim. However the sequence $(f_n)$ does not $C^0$ converge to $\mathrm{const}_0$, since this would necessarily require for $f_n$ to equal $0$ outside a compact subset for sufficiently large $n$ (see \cite{golubitsky2012stable}).
\end{remark}

Next we understand convergence in the $C^k$-topology. Before stating the main proposition we shall introduce another type of convergence which is better suited for non-compact domains, since in the non-compact case usual $C^k$-convergence is often too strong, because (as already statetd above) 
\[
f_n \xrightarrow{C^k} f
\]
would already imply that $f_n$ and $f$ agree outside a compact subset (again see \cite{golubitsky2012stable}).

\begin{definition}
    Let $X,Y$ be smooth manifolds and $f,f_n \in C^k(X,Y)$. We say that 
    \[
    f_n \xrightarrow{C^k_{\mathrm{loc}}} f
    \]
    if for every compact codimension $0$-submanifold $K \subseteq X$ we have 
    \[
    f_n|_K \xrightarrow{C^k} f|_K.
    \]
\end{definition}

Note that for compact domains $C^k_{\mathrm{loc}}$-convergence and $C^k$-convergence coincide, so we are actually interested in understanding $C^k_{\mathrm{loc}}$-convergence. The following two propositions give a useful characterization of $C^k_{\mathrm{loc}}$-convergence to work with in practice. 

\begin{proposition}
    \label{whitney_top:prop:ck_conv_on_rn}
    Let $D\subseteq \R^n$ be a compact codimension $0$ submanifold, $V \subseteq \R^m$ open and $f,f_n \in C^k(D,V)$. Then
    \[
    f_n \xrightarrow{C^k} f
    \]
    if and only if for every multiindex $\alpha \in \N^n_0$ with $0\leq\abs{\alpha} \leq k$
    \[
    \frac{\partial^{\abs{\alpha}}f_n}{\partial x^\alpha} \xrightarrow{\mathrm{uniformly}}\frac{\partial^{\abs{\alpha}}f}{\partial x^\alpha}.
    \]
\end{proposition}

\begin{proposition}
    \label{whitney_top:prop:ck_conv_equivalent_characterizations}
    Let $X,Y$ be smooth manifolds and $f,f_n \in C^k(X,Y)$. Then the following statements are equivalent.
    \begin{enumerate}[(i)]
        \item $f_n \xrightarrow{C^k_{\mathrm{loc}}} f$,
        \item $j^kf_n \xrightarrow{C^0_{\mathrm{loc}}} j^kf$,
        \item For all charts $\phi \colon U \to U^\prime \subseteq X$ and $\psi \colon V \to V^\prime \subseteq Y$ with $f(U^\prime), f_n(U^\prime) \subseteq V^\prime$ (for $n$ sufficiently large) we have 
        \[
        \psi^{-1} \circ f_n \circ \phi \xrightarrow{C^k_{\mathrm{loc}}} \psi^{-1} \circ f \circ \phi,
        \]
        \item For every $p \in X$ there exist charts $\phi \colon U \to U^\prime \subseteq X$ around $p$ and $\psi \colon V \to V^\prime \subseteq Y$ around $f(p)$ with $f(U^\prime), f_n(U^\prime) \subseteq V^\prime$ (for $n$ sufficiently large) such that
        \[
        \psi^{-1} \circ f_n \circ \phi \xrightarrow{C^k_{\mathrm{loc}}} \psi^{-1} \circ f \circ \phi.
        \]
    \end{enumerate}
\end{proposition}

Before proving the propositions we recall a few facts about the exponential topology not mentioned in \cite{steimle_topo}, which will be used in the proofs.

\begin{lemma}
    \label{whitney_top:lem:exponential_topology_facts}
    Let $X,Y,Z$ be topological spaces and $A \subseteq X$, $B \subseteq Y$ subspaces. Then the following holds (assuming the exponential topology exists on each space)
    \begin{enumerate}[(i)]
        \item The canonical map $C^0(X,B) \hookrightarrow C^0(X,Y)$ is an embedding
        \item The restriction map $C^0(X,Y) \to C^0(A,Y)$ is continuous
        \item The canonical bijection $C^0(X,Y) \times C^0(X,Z) \to C^0(X,Y \times Z)$ is a homeomorphism
    \end{enumerate}
\end{lemma}

\begin{proof}
    For (i) we show that the exponential topology on $C^0(X,B)$ agrees with the topology induced by the map  $C^0(X,B) \hookrightarrow C^0(X,Y)$ by showing the corresponding universal property. So let
    \[\begin{tikzcd}
	& {C^0(X,Y)} \\
	Z & {C^0(X,B)}
	\arrow["{F}", from=2-1, to=1-2]
	\arrow["f", from=2-1, to=2-2]
	\arrow[hook, from=2-2, to=1-2]
\end{tikzcd}\]
commute with $F$ being continuous. We have to show that $f$ is continuous, but by definiton of the exponential topology this is equivalent to 
\[
Z \times X \to B, \; (z,x) \mapsto f(z)(x)
\]
being continuous and continuity of $F$ is equivalent to continuity of 
\[
Z \times X \to Y, \; (z,x) \mapsto F(z)(x),
\]
so (i) follows. For (ii) note that the restriction map is continuous if and only if 
\[
C^0(X,Y) \times A \to Y, \; (f,x) \mapsto f|_A(x)
\]
is continuous. But this is just the restriction of the evaluation map which is continuous by \cite{steimle_topo}. Similiarly the map in (iii) is continuous if and only if 
\[
C^0(X,Y) \times C^0(X,Z) \times X \to Y \times Z, \; (f,g,x) \mapsto (f(x),g(x))
\]
is. But again the continuity follows from the continuity of the evaluation map.
\end{proof}

\begin{proof}[Proof of Proposition \ref{whitney_top:prop:ck_conv_on_rn}]
    In \cite{golubitsky2012stable} it was shown that for compact domains $C^k$-convergence is equivalent to the $k$-jets converging uniformly with respect to any metric on the Jet-bundle compatible with its topology, that is 
    \[
    f_n \xrightarrow{C^k} f \; \Leftrightarrow \; j^kf_n \xrightarrow{\mathrm{uniformly}} j^kf
    \]
    Note that for compact domains uniform convergence and $C^0_\mathrm{exp}$-convergence agree (see \cite{steimle_topo}). Now let $d(n,k) = \frac{(n+k)!}{n!k!}$ be the total number of all partial derivatives $\partial^\alpha$ of order $1 \leq \abs{\alpha} \leq k$ and let 
    \[
    \tau \colon D \times V \times \R^{d(n,k)} \xrightarrow{\cong} J^k(D,V)
    \]
    be the canonical diffeomorphism. Then 
    \[
    j^kf_n \xrightarrow{C^0_\mathrm{exp}} j^kf \; \Leftrightarrow \; j^kf_n \circ \tau \xrightarrow{C^0_\mathrm{exp}} j^kf \circ \tau.
    \]
    But a sequence of maps with values in a product $C^0_\mathrm{exp}$-converges if and only if each component $C^0_\mathrm{exp}$-converges. Since the components of $j^kf_n \circ \tau$ and $j^kf \circ \tau$ are precisely the partial derivatives of $f_n$ and $f$ up to order $k$ the statement follows (after again applying the equivalence of uniform and $C^0_\mathrm{exp}$-convergence on compact domains).
\end{proof}

\begin{proof}[Proof of Proposition \ref{whitney_top:prop:ck_conv_equivalent_characterizations}]
    Let $K \subseteq X$ be any compact codimension $0$ submanifold. It then has been shown in \cite{golubitsky2012stable} that 
    \[
    f_n|_K \xrightarrow{C^k} f|_K
    \]
    is equivalent to 
    \[
    j^kf_n|_K \xrightarrow{\mathrm{uniformly}}j^kf|_K
    \]
    with respect to any metric on $J^k(X,Y)$ compatible with the topology. But again uniform convergence on compact domains is equivalent to $C^0_\mathrm{exp}$-convergence, which in turn is equivalent to $C^0$-convergence by Lemma \ref{whitney_top:lem:c0_and_exponential_topology_agree}. This shows the equivalence of (i) and (ii). 
    
    For (ii) $\Rightarrow$ (iii) let $\phi,\psi$ be as in (iii) and let $D\subseteq U$ be any compact codimension $0$ submanifold. Then by Lemma \ref{whitney_top:lem:c0_and_exponential_topology_agree} we again observe that $C^0$-convergence of $j^kf_n$ on $\phi(D)$ is equivalent to $C^0_\mathrm{exp}$-convergence and since the exponential topology behaves nicely with respect to restiction in the codomain (see Lemma \ref{whitney_top:lem:exponential_topology_facts}) we get that the restricted maps 
    \[
    j^kf_n|_{\phi(D)} \colon \phi(D) \to J^k(U^\prime,V^\prime)
    \]
    $C^0_\mathrm{exp}$-converge to 
    \[
    j^kf|_{\phi(D)} \colon \phi(D) \to J^k(U^\prime,V^\prime),
    \]
    where $J^k(U^\prime,V^\prime) \subseteq J^k(X,Y)$ is considered an open subset. Under the diffeomorphisms
    \begin{align*}
        D &\xrightarrow{\cong} \phi(D), \\
        U \times V \times \R^{d(n,k)} &\xrightarrow{\cong} J^k(U^\prime,V^\prime)
    \end{align*}
    induced by $\phi$ and $\psi$ the maps $j^kf_n|_{\phi(D)}, j^kf|_{\phi(D)}$ correspond to maps 
    \[
    g_n,g \colon D \to U \times V \times \R^{d(n,k)}
    \]
    and $g_n \xrightarrow{C^0_\mathrm{exp}} g$. But a sequence of maps with values in a product $C^0_\mathrm{exp}$-converges if and only if each component converges (again see Lemma \ref{whitney_top:lem:exponential_topology_facts}). Since the components of $g_n$ are $\mathrm{id}_U$ and the partial derivatives of $\psi^{-1} \circ f_n \circ \phi$ of order less than $k$ (and similiarly for $g$) we conclude that 
    \[
    \frac{\partial^\alpha(\psi^{-1} \circ f_n \circ \phi)}{\partial x^\alpha}\scalebox{1}[1.8]{$|$}_D \xrightarrow{C^0_\mathrm{exp}} \frac{\partial^\alpha(\psi^{-1} \circ f \circ \phi)}{\partial x^\alpha}\scalebox{1}[1.8]{$|$}_D
    \]    
    for all $\alpha \in \N_0^n$ with $0 \leq |\alpha| \leq k$. Now (iii) follows from Proposition \ref{whitney_top:prop:ck_conv_on_rn} (and the equivalence of uniform and $C^0_\mathrm{exp}$-convergence on compact domain). 

    Clearly (iii) $\Rightarrow$ (iv), so let's assume (iv). Using the same notation and argument as above but reversed we see that 
    \[
    (\psi^{-1} \circ f_n \circ \phi)|_D \xrightarrow{C^k} (\psi^{-1} \circ f \circ \phi)|_D
    \]
    if and only if 
    \[
    j^kf_n|_{\phi(D)} \xrightarrow{C^0_\mathrm{exp}} j^kf|_{\phi(D)},
    \]
    where we consider $j^kf_n|_{\phi(D)}$ and $j^kf|_{\phi(D)}$ as maps 
    \[
    \phi(D) \to J^k(X,Y).
    \]
    Now choose any metric on $J^k(X,Y)$ compatible with the topology. Then $C^0_\mathrm{exp}$-convergence of $j^kf_n|_{\phi(D)}$ is equivalent to uniform convergence with respect to that metric. Thus we have shown that around each point $p \in X$ theres exists some neighbourhood $\phi(D)$ of $p$, such that $j^kf_n$ converges uniformly to $j^kf$ on $\phi(D)$. But any compact codimension $0$ submanifold $K\subseteq X$ is covered by finitely many such neighbourhoods and it is easily seen that this already implies uniform convergence of $j^kf_n$ to $j^kf$ on all of $K$. Then again using the equivalence of uniform convergence and $C^0_\mathrm{exp}$-convergence on compact domains as well as Lemma \ref{whitney_top:lem:c0_and_exponential_topology_agree} we conclude that 
    \[
    j^kf_n \xrightarrow{C^0_\mathrm{loc}} j^kf.
    \]
\end{proof}

\headline{An equivalent definition of the Whitney topologies}

Instead of using jet bundles one can also directly define the Whitney $C^k$-topologies on $C^k(X,Y)$ using local charts. This is for example done in \cite{hirsch2012differential}. To fix notation let us breifly recall the definiton. Suppose we are given the following data 
\begin{itemize}
    \item A family of charts $\Phi = (U_i,\phi_i)_{i\in I}$ for $X$, such that $(U_i)_{i\in I}$ is locally finite, that is each point $x \in X$ posseses a neighbourhood that interesects only finitely many $U_i$
    \item A family of charts $\Psi = (V_i,\psi_i)_{i \in I}$ for $Y$
    \item A family $K= (K_i)_{i \in I}$ of compact sets $K_i \subseteq U_i$
    \item A family of positive numbers $\epsilon = (\epsilon)_{i \in I}$
\end{itemize} 
Then if $f \in C^k(X,Y)$ with $f(K_i) \subseteq V_i$ for each $i \in I$ we define a basic open neighbourhood
\[
\mathcal{N}^k(f,\Phi,\Psi,K,\epsilon)
\]
as the set of functions $g \in C^k(X,Y)$, such that for all $i \in I$, $g(K_i) \subseteq V_i$ and 
\[
\norm{\partial^\alpha(\psi_i^{-1}f\phi_i) - \partial^\alpha(\psi_i^{-1}g\phi_i)} < \epsilon_i
\]
for all $x \in \phi^{-1}(K_i)$ and $\alpha \in \N_0^n$ with $\abs{\alpha} \leq k$ (where $n = \mathrm{dim}(X)$). The collection of all such basic open neighbourhoods form the basis for a topology on $C^k(X,Y)$ (it is easy to see that the intersection of two such neighbourhoods is again of the same form). The topology on $C^\infty(X,Y)$ is then simply defined as the topology induced by all inclusion maps $C^\infty(X,Y) \to C^k(X,Y)$ (this agrees with the definition in \cite{golubitsky2012stable}).

We now want to show that this construction does in fact yield the Whitney $C^k$-topolgies. For the moment, let us call the topology defined above the ``local chart topology'' and the topology defined in \cite{golubitsky2012stable} using jet bundles the ``jet topology''. Before showing equivalence we introduce some notation used in the proof. Suppose we are given the data $(U_i,\phi_i),(V_i,\psi_i),K_i,\epsilon_i$ for an open neighbourhood $\mathcal{N}^k(f,\Phi,\Psi,K,\epsilon)$ as explained above. We abbreviate 
\begin{align*}
    \mathcal{N}_i &:= \mathcal{N}^k(f,(U_i,\phi_i),(V_i,\psi_i),K_i,\epsilon_i), \\ \mathcal{N} &:= \mathcal{N}^k(f,\Phi,\Psi,K,\epsilon)
\end{align*}
and $ \tilde{U}_i := \phi_i^{-1}(U_i), \;
\tilde{K}_i := \phi_i^{-1}(K_i), \;
\tilde{V}_i := \psi_i^{-1}(V_i), \;
\tilde{f} := \psi_i^{-1} f \phi_i. $
Define
\[
\tilde{W}_i := \{(x,y,(v_\alpha)_{\alpha}) \in 
\tilde{U}_i \times \tilde{V}_i \times \R^{d(n,k)} 
\mid 
\norm{\tilde{f}(x) - y} < \epsilon_i, \;
\norm{\partial^\alpha \tilde{f}(x) - v_\alpha} < \epsilon_i \text{ for all } \alpha
\},
\]
where $\alpha \in \N_0^n$ ranges over all multiindices with $1 \leq \abs{\alpha} \leq k$, $v_\alpha \in \R^n$ and $d(n,k)$ is the total number of such multiindices as defined in the proof of Proposition \ref{whitney_top:prop:ck_conv_on_rn}. The charts $\phi_i$ and $\psi_i$ define a chart 
\[
\tau_i \colon \tilde{U}_i \times \tilde{V}_i \times \R^{d(n,k)} \xrightarrow{\cong} J^k(U_i,V_i) \subseteq J^k(X,Y)
\]
which we use to define 
\[
W_i := \tau(\tilde{W}_i).
\]
Clearly $W_i \subseteq J^k(X,Y)$ is open. Moreover $W_i$ is constructed in such a way that 
\begin{equation}
    \label{whitney_top:eq:eq1}
g \in \mathcal{N}_i \Leftrightarrow j^kg(K_i) \subseteq W_i.
\end{equation}

We now show that any open set in the local chart topology is also open in the jet topology. To this end consider 
\[
F_i := W_i \cup \alpha^{-1}(X \setminus K_i),
\]
where $\alpha \colon J^k(X,Y) \to X$ is the source map. Then from (\ref{whitney_top:eq:eq1}) we get 
\[
g \in \mathcal{N} _i \Leftarrow j^kg(X) \subseteq F_i \Leftrightarrow g \in \mathrm{M}(F_i)
\]
Since 
\[
\mathcal{N} = \bigcap_{i \in I} \mathcal{N} _i
\]
we can show that $\mathcal{N} \subseteq C^k(X,Y)$ is open in the jet topology if we can show that 
\[
F := \bigcap_{i \in I} F_i \subseteq J^k(X,Y)
\] 
is open, because $\bigcap_{i \in I} \mathrm{M}(F_i) = \mathrm{M}(F)$. For this we will use the local finiteness of the sets $U_i$. Let $\sigma \in W$ be a jet. Then around its source $\alpha(\sigma) \in X$ we can find an open neighbourhood
\[
A \subseteq X,
\]
such that $A$ only intersects finitely many (possibly none) of the $U_i$, say $U_{i_1}, \ldots, U_{i_N}$. Then 
\[
B:= W_{i_1} \cap \ldots \cap W_{i_N} \cap \alpha^{-1}(A) \subseteq J^k(X,Y) 
\] 
is open and for any $\sigma^\prime \in B$ we have 
\[
\alpha(\sigma^\prime) \in A \subseteq X \setminus K_i 
\]
for all $i \neq i_1,\ldots,i_N$, thus $\sigma \in F_i$ for all $i \in I$ and hence $F$ is open.

Now we show that for an open set $T \subseteq J^k(X,Y)$ the set $\mathrm{M}(T) \subseteq C^k(X,Y)$ is also open in the local chart topology. For $f \in \mathrm{M}(T)$ choose charts $(\phi_i,U_i)$ of $X$, $(\psi_i,V_i)$ of $Y$ and compact sets $K_i \subseteq U_i$, such that the $U_i$ are a locally finite open cover of $X$, the sets $K_i$ still cover $X$ and $f(K_i) \subseteq V_i$ for all $i \in I$. We have to find positive numbers $\epsilon_i > 0$, such that the basic open neighbourhood $\mathcal{N}$ defined using this data is contained in $\mathrm{M}(T)$. Suppose that for each $i \in I$ we can find $\epsilon_i >0$, such that 
\begin{equation}
    \label{whitney_top:eq:eq2}
    \tilde{W}_i \cap (\tilde{K}_i \times \tilde{V}_i \times \R^{d(n,k)}) \subseteq \tau_i^{-1}(T) \cap (\tilde{K}_i \times \tilde{V}_i \times \R^{d(n,k)}).
\end{equation}
For any $g \in \mathcal{N}$ we have 
\[
(\tau^{-1}\circ j^kg)(K_i) \subseteq \tilde{W}_i \cap (\tilde{K}_i \times \tilde{V}_i \times \R^{d(n,k)}) 
\]
and thus we get from (\ref{whitney_top:eq:eq2})
\[
j^kg(K_i) \subseteq T.
\] 
This holds for all $i \in I$ and since the $K_i$ cover $X$ we conclude that 
\[
j^kg(X) \subseteq T.
\]
Thus $\mathcal{N} \subseteq \mathrm{M}(T)$. To show the existence of these $\epsilon_i$ we invoke the following Lemma from point set topology.

\begin{lemma}
    Let $D$ be a compact topological space, $(Z,d)$ a metric space and $h \colon D \to Z$ continuous. Then for any open neighbourhood $S \subseteq D \times Z$ of the graph $\Gamma_h \subseteq D \times Z$ of $h$, there exists $\epsilon >0$, such that the $\epsilon$-tube 
    \[
    \{
        (x,z) \in D \times Z \mid d(h(x),z) < \epsilon
    \}
    \]
    is contained in $S$.
\end{lemma}

\begin{proof}
    See Lemma \ref{point_set_topo:lem:tube_lemma}.
\end{proof}
\noindent Applying this Lemma to 
\begin{itemize}
\item $D = \tilde{K}_i$,
\item $Z=\tilde{V}_i \times \R^{d(n,k)}$,
\item $h = (\tilde{f}, (\partial^\alpha\tilde{f})_\alpha)$,
\item $S = \tau_i^{-1}(T) \cap (\tilde{K}_i \times \tilde{V}_i \times \R^{d(n,k)})$
\end{itemize}
with the correct norm on $\R^{d(n,k)}$ yields the desired $\epsilon_i$.





