\subsection{Chern and Pontryagin Classes}

    Let $E \to M$ be a complex vector bundle of rank $d$. For any matrix $A \in \C^{d \times d}$ we can consider the characteristic polynomial in $t$:
    \[
    \mathrm{det}\left( \mathrm{id} - \frac{t}{2\pi i} A\right) = \sum_{k=0}^{d} f_k(A)t^k,
    \]
    where $f_k \colon \C^{d \times d} \to \C$ are the coefficients of this polynomial and depend on $A$. Using the Leibniz formula for the determinant one can see that $f_k$ is a $k$-homogeneous polynomial and it moreover is invariant since the characteristic polynomial is. Furthermore it follows from the Leibniz formula that 
    \begin{align*}
        f_0(A) & = 1, \\
        f_1(A) &= \mathrm{tr}\left(-\frac{1}{2\pi i} A\right), \\
        f_d(A) &= \mathrm{det}\left(-\frac{1}{2\pi i} A\right).
    \end{align*}

\begin{definition}
    The \textbf{$k$-th Chern class} of $E$ is defined as 
    \[
    c_k(E) := \chi_E(f_k) \in H_\mathrm{dR}^{2k}(M;\C).
    \]
\end{definition}

If we instead consider a real vector bundle $E \to M$ we can similiarly define $k$-homogeneous polynomials $g_k \colon \R^{d \times d} \to \R$ by 
\[
    \mathrm{det}\left( \mathrm{id} - \frac{t}{2\pi} A\right) = \sum_{k=0}^{d} g_k(A)t^k,
\]
and define:

\begin{definition}
    The \textbf{$k$-th Pontryagin class} of $E$ is defined as 
    \[
    p_k(E) := \chi_E(g_{2k}) \in H_\mathrm{dR}^{4k}(M;\R).
    \]
\end{definition}

Of course there is a connection between these two definitions:

\begin{lemma}
    Let $E \to M$ be a real vector bundle. Then 
    \[
    p_{k}(E) = (-1)^k c_{2k}(E \otimes \C),
    \]
    where $E \otimes \C$ is the complexified bundle and the equality is understood by considering a real differential as the real part of a complex differential form, that is $H_\mathrm{dR}^{4k}(M;\R) \subseteq H_\mathrm{dR}^{4k}(M;\C)$.
\end{lemma}

\begin{proof}
    Choose any connection $\nabla$ on $E$ and let $\nabla_\C$ be the induced connection on the complexified bundle $E \otimes \C$. Then by definition any frame of $E$ also defines a frame of $E \otimes \C$ and the corresponding connection $1$-forms are equal. Thus the curvature form $\Omega$ of $E$ agrees with the curvature form $\Omega_\C$ of $E \otimes \C$. From the definition of the polynomials $f_k$ and $g_k$ it is moreover clear that for any real matrix $A \in \R^{d \times d}$ we have 
    \[
    f_k(A) = i^k g_k(A).
    \]
    Thus 
    \[
    c_{2k}(E) = [f_{2k}(\Omega_\C)] = [f_{2k}(\Omega)] = i^{2k} [g_{2k}(\Omega)] = (-1)^k p_k(E).
    \]
\end{proof}

Now assume we are given a real oriented vector bundle $E \to M$ of rank $2n$. Choose a metric on $E$ together with a compatible connection. A simple computation shows that the connection $1$-form $\Theta$ with respect to any oriented orthonormal frame is skew-symmetric. Thus 
\begin{align*}
    (\Theta \wedge \Theta)^T(X,Y) &= (\Theta(X)\Theta(Y) - \Theta(Y)\Theta(X))^T \\ 
    &= (-\Theta(Y))(-\Theta(X)) - (-\Theta(X))(-\Theta(Y)) \\
    &= - (\Theta \wedge \Theta) (X,Y).
\end{align*}
So the curvature form $\Omega$ is also skew symmetric:
\[
\Omega^T = (d\Theta)^T + (\Theta \wedge \Theta)^T = -d\Theta - \Theta \wedge \Theta = -\Omega.
\]
Recall that the Pfaffian is a $n$-homogeneous polynomial 
\[
\mathrm{Pf} \colon \mathrm{Skew}_{2n} \to \R
\]
on the space of skew symmetric $2n \times 2n$ matrices and satisifies 
\[
\mathrm{Pf}(B^T A B) = \mathrm{det}(B)\mathrm{Pf}(A)
\]
for $A \in \mathrm{Skew}_{2n}$ and $B \in \R^{n \times n}$ an arbitrary matrix. Thus we can define the $2n$-form $\mathrm{Pf}(\Omega)$ on the neighbourhood where the oriented orthonormal frame was defined. Under a change of oriented orthonormal frame the curvature form $\Omega$ transforms as $S \Omega S^{-1}$ for a special orthogonal matrix $S \in \mathrm{SO}(2n)$. Since $S^{-1} = S^T$ and $\mathrm{det}(S) = 1$ we conclude that $\mathrm{Pf}(\Omega)$ does not depend on the chosen frame and thus defines a global object. Just as in the proof of Theorem \ref{chern_weil:thm:chern_weil_thm} we see that $\mathrm{Pf}(\Omega)$ is closed (only the statement of Lemma \ref{chern_weil:lem:technical_lemma1} changes, where instead of arbitrary $X_i$ and $Y$ we only allow for skew symmetric matrices. Note that the tangent space of $\mathrm{SO}(2n)$ is precisely the space of skew symmetric matrices, such that we can realize any such $Y$ as the velocity vector of a curve through $\mathrm{id}$). Moreover the cohomology class does not depend on the chosen connection and metric. Again the proof is similiar to Theorem \ref{chern_weil:thm:chern_weil_thm}. Thus we can define 

\begin{definition}
    The \textbf{Euler class} of $E$ is defined as 
    \[
    e(E) := \left[\mathrm{Pf}\left(\frac{1}{2\pi}\Omega\right)\right] \in H_\mathrm{dR}^{2n}(M;\R).
    \]
\end{definition}