\subsection{Smoothness on arbitrary subsets}

Let $M,N$ be manifolds and $A \subseteq M$ an arbitrary subset. In \cite{lee2011smooth} the author defines what it means for a map
\[
f \colon A \to N
\]
to be smooth, namely
\begin{center}
    \textit{We call $f$ smooth if around each $x \in A$ there exists a smooth extension $\oset[-1pt]{\sim}{f} \colon U \to N$ of $f$ to some open neighbourhood $U \subseteq M$ of $x$ (see \cite[p. 45]{lee2011smooth}).}
\end{center}
However this definition has a fatal flaw: if $A$ happens to be a submanifold of $M$, a map $f \colon A \to N$ might be smooth in the usual sense (i.e. as a map between smooth manifolds) but not smooth in the sense of Lee`s definition. Consider for example
\begin{counterexample}
    \label{ex:extension_smooth}
    \begin{gather*}
        M := \R, N:= [0,1], A := [0,1], \\
        f \colon A \to N, \, t \mapsto t.   
    \end{gather*}
    Then clearly $f$ is smooth as a map between smooth manifolds, but around $0,1 \in A$ it does not admit a smooth extension onto an open neighbourhood $U \subseteq M$, since the image of such an extension would fail to be contained within $N$. 
\end{counterexample}
We can however correct this flaw by defining:

\begin{definition}[Extension-smooth]
    \label{def:extension_smooth}
    Let $M^n,N^m$ be manifolds with or without boundary, $A \subseteq M$ an arbitrary subset and $f \colon A \to N$ a map. We call $f$ \textit{extension-smooth} if around each $x \in A$ there exist charts
    \begin{gather*}
        h \colon U \to M \text{ around } x\\
        k \colon V \to N \text{ around } f(x),
    \end{gather*}
    such that
    \[
    g:= k^{-1} \circ f \circ h|_{h^{-1}(A)}
    \]
    admits a smooth extension around $h^{-1}(x)$ in the sense that there exist open subsets $U_0 \subseteq \R^n$ and $V_0 \subseteq \R^m$ with $h^{-1}(x) \in U_0$ and $k^{-1}(f(x)) \in V_0$ and a smooth map 
    \[
    \oset[-1pt]{\sim}{g} \colon U_0 \to V_0
    \]
    with $g|_{U_0 \cap h^{-1}(A)} = \oset[-1pt]{\sim}{g}|_{U_0 \cap h^{-1}(A)}$.
\end{definition}

We now claim the following.

\begin{lemma}
    \label{lem:smoothness_arbitrary_subsets_1}
    If $\partial N = \emptyset$ then Lee`s definition and Definition \ref{def:extension_smooth} are equivalent. In general, if $f$ is smooth in the sense of Lee`s definition, then it is also extension-smooth but the converse may fail.
\end{lemma}

\begin{lemma}
    If $A$ is a submanifold of $M$, then $f \colon A \to N$ is smooth as a map of smooth manifolds if and only if it is extension-smooth.
\end{lemma}

In particular the extension Lemma in \cite{lee2011smooth} (Lemma 2.26.) still holds and the proof of the Whitney approximation theorems in chapter 6 work the same (here the extension Lemma is used). But with our new definition Problem 6-7 in \cite{lee2011smooth} now is an actual counterexample. (with Lee`s definition $F$ would not be smooth on the closed subset $A$ but only smooth as a map between manifolds).

Another way to properly define smoothness of $f \colon A \to N$ would be to change the codomain in Lee`s definition to be the double of $N$. This should be equivalent to Definition \ref{def:extension_smooth} and is arguably more elegant at the cost of being less elementary. 

\headline{Smooth Homotopies}

\noindent This allows to define a \textbf{smooth homotopy} as a smooth map 
\[
H \colon M \times [0,1] \to \N,
\]
for manifolds $M,N$ with or without boundary, where we consider $N \times [0,1]$ as a subset of $N \times \R$ (in the sense of Definition \ref{def:extension_smooth}). Since the definition of extension-smooth is not that nice to work with we may instead want to define:
\begin{definition}
    \label{def:smooth_homotopy}
    A homotopy $H\colon M \times [0,1] \to N$ is called smooth if there exists a smooth extension of $H$ to $M \times \R$.
\end{definition}
As we have seen in Counterexample \ref{ex:extension_smooth} this is a stronger definition compared to the first one given, as less homotopies are considered to be smooth by Definition \ref{def:smooth_homotopy}. However the homotopy classes of smooth maps are the same. More precisely this is formulated in the following lemma.

\begin{lemma}
    Let $M,N$ be manifolds with or without boundary and $f,g \colon M \to N$ smooth maps. Then $f$ and $g$ are smoothly homotopic in the sense of the first definition if and only if they are smoothly homotopic in the sense of Definition \ref{def:smooth_homotopy}.
\end{lemma}

\begin{proof}[Proof]
    Let $H \colon M \times [0,1] \to N$ be a smooth homotopy from $f$ to $g$ in the sense of the first definition (i.e. it is extension-smooth). Choose a smooth function $\phi \colon \R \to [0,1]$ with
    \begin{gather*}
        \phi(t) = 0 \text{ for } t \leq 0 \\
        \phi(t) = 1 \text{ for } t \geq 1
    \end{gather*}
    Then 
    \[
        \Phi \colon M \times [0,1] \to M \times [0,1], (x,t) \mapsto (x,\phi(t))
    \]
    is extension-smooth (it can be smoothly extended to $M \times \R$ and this implies that it is smooth in the sense of Definition \ref{def:extension_smooth}, as Lee`s definition is stronger by Lemma \ref{lem:smoothness_arbitrary_subsets_1}). Furthermore, it is easily verified that the composition of extension-smooth maps on arbitray subsets is extension-smooth, such that 
    \[
    \overline{H} := H \circ \Phi \colon M \times [0,1] \to N
    \] 
    is extension-smooth. Then $\overline{H}$ is a homotopy from $f$ to $g$ and we can extend $\overline{H}$ to the smooth map 
    \[
        G \colon M \times \R \to N, (x,t) \mapsto H(x,\phi(t)),
    \]
    hence $\overline{H}$ is a smooth homotopy from $f$ to $g$ in the sense of Definition \ref{def:smooth_homotopy}. The converse follows directly from Lemma \ref{lem:smoothness_arbitrary_subsets_1}.
    
\end{proof}


