\subsection{Formulas for the Hodge star operator and codifferential}

In this section we briefly discuss the definition of the Hodge star operator and codifferential as well as their basic properties. This is supposed to serve as a canonical reference for the correct signs in the formulas involved. After the basic properties we proof a few more formulas that might be interesting. 

\par

Let $(V,\langle\cdot,\cdot\rangle)$ be an oriented $n$-dimensional euclidian vector space. Then the Hodge star operator is the unique linear map 
\[
\star \colon \Lambda^kV \to \Lambda^{n-k}V
\]
satisfying 
\[
\alpha \wedge \star \beta = \langle\alpha,\beta\rangle \mathrm{vol},
\]
for all $\alpha,\beta \in \Lambda^k V$, where $\mathrm{vol} \in \Lambda^nV$ is the canonical volume form on $V$. If $e_1,\ldots,e_n \in V$ is an oriented orthonormal basis we see that for any ordered tuple $(i_1,\ldots,i_k)$
\[
\star(e_{i_1} \wedge \ldots \wedge e_{i_k}) = e_{j_1} \wedge \ldots \wedge e_{j_{n-k}},
\]
where $(j_1,\ldots,j_{n-k})$ is the complementary ordered tuple. Since tuples of this form define an orthonormal basis it follows that $\star$ is an isometry. Using this fact we compute
\begin{align*}
    \langle\star\star\beta,\alpha\rangle \mathrm{vol} &= \star\star\beta\wedge\star\alpha = (-1)^{k(n-k)}\star\alpha\wedge\star\star\beta \\ 
    &= (-1)^{k(n-k)}\langle\star\alpha,\star\beta\rangle\mathrm{vol}
    = (-1)^{k(n-k)}\langle\alpha,\beta\rangle\mathrm{vol}
\end{align*}
from which we conclude that 
\[
\star\star = (-1)^{k(n-k)} \mathrm{id}_{\Lambda^kV}.
\]
The Hodge star operator can also be used to relate the interior and exterior product. For $v \in V$ and $\omega \in \Lambda^k V^\ast$ we denote the interior product by $v \intprod \omega \in \Lambda^{k-1}V^\ast$, that is
\[
(v \intprod \omega )(v_1,\ldots,v_{k-1}) := \omega(v,v_1,\ldots,v_{k-1}).
\]
(Some authors define this with the opposite sign, i.e. $v \intprod \omega := - \omega(v,-)$). Moreover the notation $v \wedge \omega \in \Lambda^{k+1}V^\ast$ is to be understood as first dualizing $v$ and then taking the usual wedge product, that is 
\[
v \wedge \omega := \langle v,-\rangle\wedge\omega.
\]
We then have the following Lemma.

\begin{lemma}
    \label{cartan_calc:lem:interior_product_and_hodge_star}
    For $v\in V$, $\omega \in \Lambda^k V$ and $\eta \in \Lambda^{k-1}V$ the following holds 
    \begin{enumerate}[(i)]
        \item $\langle \eta, v \intprod \omega \rangle = \langle v \wedge\eta, \omega \rangle$
        \item $v \intprod \omega = (-1)^{n(k+1)} \star (v \wedge \star \omega)$
    \end{enumerate}
\end{lemma}

\begin{proof}
    For $(i)$ note that the expression on both sides is linear in $v, \omega$ and $\eta$, such that it suffices to show equality on basis vectors. So let $e_1,\ldots,e_n \in V$ be an oriented orthonormal basis and $\epsilon^1,\ldots,\epsilon^n \in V^\ast$ its dual basis. We have to compute $(i)$ for
    \[
    v = e_m, \;\; \omega = \epsilon^{i_1} \wedge \ldots \wedge \epsilon^{i_k}, \;\; \eta = \epsilon^{j_1} \wedge \ldots \wedge \epsilon^{j_{k-1}}.
    \]
    First note that we may assume $m = i_p$ and
    \[
    (i_1,\ldots,\widehat{i_p},\ldots,i_k) = (j_1,\ldots,j_{k-1}).
    \] 
    for some $1 \leq p \leq k$ since otherwise both sides vanish. But then
    \[
    v \intprod \omega = (-1)^{p-1} \eta
    \]
    and 
    \[
    v \wedge \eta = (-1)^{p-1} \omega
    \]
    from which $(i)$ follows. Using this we can show $(ii)$:
    \begin{align*}
        \langle \eta , v \intprod \omega \rangle \mathrm{vol} & = \langle v \wedge \eta, \omega \rangle \mathrm{vol}\\
        &= v \wedge \eta \wedge \star \omega \\
        &= (-1)^{k-1} \eta \wedge v \wedge \star \omega \\
        &= (-1)^{k-1 + (n-k+1)(k-1)} \eta \wedge \star\star(v \wedge \star \omega) \\
        &= (-1)^{(n-k+2)(k-1)} \langle \eta, \star (v \wedge \star \omega)\rangle \mathrm{vol}
    \end{align*}
    Moreover we compute modulo $2$:
    \[
        (n-k + 2) (k-1) = (n-k)(k-1) = n(k-1) - k(k-1) = n(k+1)
    \]
    Since $\eta$ was arbitrary the statement follows.
\end{proof}

\begin{remark}
    Part $(i)$ of Lemma \ref{cartan_calc:lem:interior_product_and_hodge_star} states that the interior product is the adjoint of the exterior product.
\end{remark}

Next we show a useful formula for the exterior derivative and codifferential in terms of the Levi-Civita connection. Let $(M,g)$ be an oriented $n$-dimensional Riemannian manfiold. First let's fix a sign convention for the codifferential $\delta$. On $k$-forms we define $\delta$ by 
\[
\begin{tikzcd}
	{\Omega^k(M)} & {\Omega^{k-1}(M)} \\
	{\Omega^{n-k}(M)} & {\Omega^{n-k+1}(M)}
	\arrow["{\textstyle (-1)^k\delta}", from=1-1, to=1-2]
	\arrow["{\textstyle \star}", from=1-1, to=2-1]
	\arrow["{\textstyle \star}", from=1-2, to=2-2]
	\arrow["{\textstyle d}", from=2-1, to=2-2]
\end{tikzcd}
\]
or in formulas 
\[
\delta := (-1)^k\star^{-1} d \star = (-1)^{n(k-1)+1} \star d \star \colon \Omega^k(M) \to \Omega^{k-1}(M).
\]
With this sign convention we have for $\omega \in \Omega^k(M)$ and $\eta \in \Omega^{k+1}(M)$
\begin{equation}
    d(\omega \wedge \star \eta) = d\omega \wedge \star\eta - \omega \wedge \star \delta \eta. \label{cartan_calc:eq:product_rule_with_codifferential}
\end{equation}
If $\alpha,\beta \in \Omega^k(M)$, such that the intersection of their supports is compact, we denote by 
\[
\langle\langle \alpha,\beta \rangle\rangle := \int_M \langle \alpha,\beta \rangle \mathrm{vol}
\]
their $L^2$-scalar product. Using Stoke's Theorem it then follows from (\ref{cartan_calc:eq:product_rule_with_codifferential}):

\begin{lemma}
    Let $\omega \in \Omega^k(M)$ and $\eta \in \Omega^{k+1}(M)$, such that the intersection of their supports is compact and contained in $M \setminus \partial M$. Then 
    \[
    \langle\langle d\omega,\eta \rangle\rangle = \langle\langle \omega,\delta\eta \rangle\rangle.
    \]
\end{lemma}

With these sign conventions in mind we can now show:

\begin{lemma}
    Let $M$ be a Riemannian manifold and $\omega \in \Omega^\ast(M)$. Let $e_1,\ldots,e_n$ be an orthonormal local frame of $TM$, then 
    \[
    d\omega = \sum_{i=1}^n e_i \wedge \nabla_i\omega,
    \]
    where $\nabla_i \omega := \nabla_{e_i}\omega$ denotes the connection on the bundle $\Lambda^\ast T^\ast M$ induced by the Levi-Civita connection on $M$. Moreover, if $M$ is oriented we have 
    \[
    \delta\omega =  - \sum_{i=1}^n e_i \intprod \nabla_i\omega.
    \]
\end{lemma}

\begin{proof}
    We first show the formula for the exterior derivative for $1$-forms $\alpha \in \Omega^1(M)$. Using Corollary \ref{cartan_cal:cor:formula_d_alpha} and the symmetry of the Levi-Civita connection we get
    \begin{align*}
        d\alpha(X,Y) &= X \cdot \alpha(Y) - Y \cdot \alpha(X) - \alpha([X,Y]) \\
        &= \nabla_X(\alpha(Y)) - \nabla_Y(\alpha(X)) - \alpha(\nabla_XY - \nabla_YX) \\
        &= (\nabla_X\alpha)(Y) - (\nabla_Y\alpha)(X)
    \end{align*}
    for any smooth vector fields $X,Y \in \mathscr{X}(M)$. Let $\epsilon^1,\ldots,\epsilon^n$ be the dual frame of $e_1,\ldots,e_n$. It then follows that 
    \begin{align*}
        d\alpha &= \sum_{i,j=1}^n d\alpha(e_i,e_j) \epsilon^i \otimes \epsilon^j \\ 
        &= \sum_{i=1}^{n} \epsilon^i \otimes \left(\sum_{j=1}^{n} (\nabla_i\alpha)(e_j)\epsilon^j\right) - \sum_{j=1}^{n} \left(\sum_{i=1}^{n} (\nabla_j\alpha)(e_i)\epsilon^i\right) \otimes \epsilon^j \\
        &= \sum_{i=1}^{n} \epsilon^i \otimes \nabla_i\alpha - \sum_{j=1}^{n} \nabla_j\alpha \otimes \epsilon^j \\ 
        &= \sum_{i=1}^n \epsilon^i \wedge \nabla_i\alpha \\
        &=  \sum_{i=1}^n e_i \wedge \nabla_i\alpha
    \end{align*}
    where we have used $\epsilon^i = \langle e_i,- \rangle$ in the last equality (since $e_1,\ldots,e_n$ is orthonormal). Now define a differential operator $D$ by 
    \[
    D\omega := \sum_{i=1}^n e_i \wedge \nabla_i\omega.
    \]
    Clearly $D$ agrees with the exterior derivative on $0$-forms and we have shown above that they furthermore agree on $1$-forms. Moreover
    \begin{align*}
        D(\omega\wedge\eta) &= \sum_{i=}^{n} e_i \wedge \nabla_i\omega \wedge \eta + e_i \wedge \omega \wedge\nabla_i\eta \\
        &= (D\omega)\wedge\eta + (-1)^{\abs{\omega}}\omega \wedge D\eta.
    \end{align*}
    So $D$ and $d$ also satisfy the same Leibniz rule. It follows that $D=d$. From this we can deduce the formula for the codifferential by using Lemma \ref{cartan_calc:lem:interior_product_and_hodge_star}. So let $\omega \in \Omega^k(M)$.
    \begin{align*}
        \delta \omega &= (-1)^{n(k-1)+1} \star d \star \omega\\
        &= (-1)^{n(k-1)+1} \sum_{i=1}^n \star (e_i \wedge \nabla_i\star\omega) \\
        &=(-1)^{n(k-1)+1} \sum_{i=1}^n \star (e_i \wedge \star\nabla_i\omega) \\
        &=(-1)^{n(k-1)+1 + n(k+1)} \sum_{i=1}^{n} e_i \intprod \nabla_i\omega \\
        &= -\sum_{i=1}^{n} e_i \intprod \nabla_i\omega.
    \end{align*}
    In the third equality we used the fact that covariant differentiation and the Hodge star operator commute, as shown in the Lemma \ref{cartan_cal:lem:hodge_star_and_connection_commute} below.
\end{proof}


\begin{lemma}
    \label{cartan_cal:lem:hodge_star_and_connection_commute}
    Let $M$ be a an oriented Riemannian manifold, $\omega \in \Omega^\ast(M)$ and $X \in \mathscr{X}(M)$. Then 
    \[
    \nabla_X\star\omega=\star\nabla_X\omega.
    \]
\end{lemma}
\begin{proof}
    Let $\omega \in \Omega^k(M)$ and $\eta \in \Omega^{n-k}(M)$. Recall that the volume form $\mathrm{vol} \in \Omega^n(M)$ is parallel, i.e. $\nabla_X\mathrm{vol} = 0$. Hence 
    \begin{align*}
        \nabla_X(\langle\eta,\star\omega\rangle\mathrm{vol}) = (\langle\nabla_X\eta,\star\omega\rangle + \langle\eta,\nabla_X\star\omega\rangle)\mathrm{vol}
    \end{align*}
    On the other hand 
    \begin{align*}
        \nabla_X(\langle\eta,\star\omega\rangle\mathrm{vol}) &= \nabla_X(\eta \wedge \star\star\omega) \\
        &= (-1)^{k(n-k)}(\nabla_X\eta \wedge \omega + \eta \wedge \nabla_X\omega) \\
        &=\nabla_X\eta \wedge \star\star\omega + \eta \wedge \star\star\nabla_X\omega \\
        &= (\langle\nabla_X\eta,\star\omega\rangle + \langle\eta,\star\nabla_X\omega\rangle)\mathrm{vol}.
    \end{align*}
    Thus 
    \[
    \langle\eta,\nabla_X\star\omega\rangle = \langle\eta,\star\nabla_X\omega\rangle.
    \]
    Since $\eta$ was arbitrary the statement follows.
\end{proof}
