\subsection{An equivalent definition of the Whitney Toplogies}
\label{whitney_top:sec:equivalent_def}

In this section we want to show that the Whitney $C^k$-topologies on $C^k(X,Y)$ as defined at the beginning of section \ref{whitney_top:sec:approx_results} agree with the definition given in \cite{golubitsky2012stable} based on jet bundles.

For the moment, let us call the topology defined in section \ref{whitney_top:sec:approx_results} the ``local chart topology'' and the topology defined in \cite{golubitsky2012stable} using jet bundles the ``jet topology''. First note that in both cases the topology on $C^\infty(X,Y)$ is defined as the topology induced by all inclusion maps $C^\infty(X,Y) \to C^k(X,Y)$, so in the following we assume $k$ to be finite. Before showing equivalence we introduce some notation used in the proof. Suppose we are given the data $(U_i,\phi_i),(V_i,\psi_i),K_i,\epsilon_i$ for an open neighbourhood $\mathcal{N}^k(f,\Phi,\Psi,K,\epsilon)$ as explained in section \ref{whitney_top:sec:approx_results}. We abbreviate 
\begin{align*}
    \mathcal{N}_i &:= \mathcal{N}^k(f,(U_i,\phi_i),(V_i,\psi_i),K_i,\epsilon_i), \\ \mathcal{N} &:= \mathcal{N}^k(f,\Phi,\Psi,K,\epsilon)
\end{align*}
and $ \tilde{U}_i := \phi_i^{-1}(U_i), \;
\tilde{K}_i := \phi_i^{-1}(K_i), \;
\tilde{V}_i := \psi_i^{-1}(V_i), \;
\tilde{f} := \psi_i^{-1} f \phi_i. $
Define
\[
\tilde{W}_i := \{(x,y,(v_\alpha)_{\alpha}) \in 
\tilde{U}_i \times \tilde{V}_i \times \R^{d(n,k)} 
\mid 
\norm{\tilde{f}(x) - y} < \epsilon_i, \;
\norm{\partial^\alpha \tilde{f}(x) - v_\alpha} < \epsilon_i \text{ for all } \alpha
\},
\]
where $\alpha \in \N_0^n$ ranges over all multiindices with $1 \leq \abs{\alpha} \leq k$, $v_\alpha \in \R^n$ and $d(n,k)$ is the total number of such multiindices as defined in the proof of Proposition \ref{whitney_top:prop:ck_conv_on_rn}. The charts $\phi_i$ and $\psi_i$ define a chart 
\[
\tau_i \colon \tilde{U}_i \times \tilde{V}_i \times \R^{d(n,k)} \xrightarrow{\cong} J^k(U_i,V_i) \subseteq J^k(X,Y)
\]
which we use to define 
\[
W_i := \tau(\tilde{W}_i).
\]
Clearly $W_i \subseteq J^k(X,Y)$ is open. Moreover $W_i$ is constructed in such a way that 
\begin{equation}
    \label{whitney_top:eq:eq1}
g \in \mathcal{N}_i \Leftrightarrow j^kg(K_i) \subseteq W_i.
\end{equation}

We now show that any open set in the local chart topology is also open in the jet topology. To this end consider 
\[
F_i := W_i \cup \alpha^{-1}(X \setminus K_i),
\]
where $\alpha \colon J^k(X,Y) \to X$ is the source map. Then from (\ref{whitney_top:eq:eq1}) we get 
\[
g \in \mathcal{N} _i \Leftarrow j^kg(X) \subseteq F_i \Leftrightarrow g \in \mathrm{M}(F_i)
\]
Since 
\[
\mathcal{N} = \bigcap_{i \in I} \mathcal{N} _i
\]
we can show that $\mathcal{N} \subseteq C^k(X,Y)$ is open in the jet topology if we can show that 
\[
F := \bigcap_{i \in I} F_i \subseteq J^k(X,Y)
\] 
is open, because $\bigcap_{i \in I} \mathrm{M}(F_i) = \mathrm{M}(F)$. For this we will use the local finiteness of the sets $U_i$. Let $\sigma \in W$ be a jet. Then around its source $\alpha(\sigma) \in X$ we can find an open neighbourhood
\[
A \subseteq X,
\]
such that $A$ only intersects finitely many (possibly none) of the $U_i$, say $U_{i_1}, \ldots, U_{i_N}$. Then 
\[
B:= W_{i_1} \cap \ldots \cap W_{i_N} \cap \alpha^{-1}(A) \subseteq J^k(X,Y) 
\] 
is open and for any $\sigma^\prime \in B$ we have 
\[
\alpha(\sigma^\prime) \in A \subseteq X \setminus K_i 
\]
for all $i \neq i_1,\ldots,i_N$, thus $\sigma \in F_i$ for all $i \in I$ and hence $F$ is open.

Now we show that for an open set $T \subseteq J^k(X,Y)$ the set $\mathrm{M}(T) \subseteq C^k(X,Y)$ is also open in the local chart topology. For $f \in \mathrm{M}(T)$ choose charts $(\phi_i,U_i)$ of $X$, $(\psi_i,V_i)$ of $Y$ and compact sets $K_i \subseteq U_i$, such that the $U_i$ are a locally finite open cover of $X$, the sets $K_i$ still cover $X$ and $f(K_i) \subseteq V_i$ for all $i \in I$. We have to find positive numbers $\epsilon_i > 0$, such that the basic open neighbourhood $\mathcal{N}$ defined using this data is contained in $\mathrm{M}(T)$. Suppose that for each $i \in I$ we can find $\epsilon_i >0$, such that 
\begin{equation}
    \label{whitney_top:eq:eq2}
    \tilde{W}_i \cap (\tilde{K}_i \times \tilde{V}_i \times \R^{d(n,k)}) \subseteq \tau_i^{-1}(T) \cap (\tilde{K}_i \times \tilde{V}_i \times \R^{d(n,k)}).
\end{equation}
For any $g \in \mathcal{N}$ we have 
\[
(\tau^{-1}\circ j^kg)(K_i) \subseteq \tilde{W}_i \cap (\tilde{K}_i \times \tilde{V}_i \times \R^{d(n,k)}) 
\]
and thus we get from (\ref{whitney_top:eq:eq2})
\[
j^kg(K_i) \subseteq T.
\] 
This holds for all $i \in I$ and since the $K_i$ cover $X$ we conclude that 
\[
j^kg(X) \subseteq T.
\]
Thus $\mathcal{N} \subseteq \mathrm{M}(T)$. To show the existence of these $\epsilon_i$ we invoke the following Lemma from point set topology.

\begin{lemma}
    Let $D$ be a compact topological space, $(Z,d)$ a metric space and $h \colon D \to Z$ continuous. Then for any open neighbourhood $S \subseteq D \times Z$ of the graph $\Gamma_h \subseteq D \times Z$ of $h$, there exists $\epsilon >0$, such that the $\epsilon$-tube 
    \[
    \{
        (x,z) \in D \times Z \mid d(h(x),z) < \epsilon
    \}
    \]
    is contained in $S$.
\end{lemma}

\begin{proof}
    See Lemma \ref{point_set_topo:lem:tube_lemma}.
\end{proof}
\noindent Applying this Lemma to 
\begin{itemize}
\item $D = \tilde{K}_i$,
\item $Z=\tilde{V}_i \times \R^{d(n,k)}$,
\item $h = (\tilde{f}, (\partial^\alpha\tilde{f})_\alpha)$,
\item $S = \tau_i^{-1}(T) \cap (\tilde{K}_i \times \tilde{V}_i \times \R^{d(n,k)})$
\end{itemize}
with the correct norm on $\R^{d(n,k)}$ yields the desired $\epsilon_i$.